---
title: "Visualizing Models in R"
subtitle: "Soc 505"
author: "Connor Gilroy"
date: "`r Sys.Date()`"
output: 
  html_document: 
    theme: cosmo
    highlight: pygments
    css: styles.css
    fig_width: 6
    toc: true
    toc_float: true
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.align = "center")
```

Today, I'll talk about working with model objects and visualizing models. 

I'll use 3 main packages: 

- `broom`
- `ggplot2`
- `ggeffects`

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
library(broom)
library(ggplot2)
library(ggeffects)

# I have aesthetic opinions
theme_set(theme_minimal())
```

# Linear regression

Fit a linear model: 

```{r}
# data from hw2
courtpenalties <- read_dta("data/courtpenalties2020_compatible.dta")

f1 <- penalties ~ minority + liberal + popsize + crt + unemp
model1 <- glm(f1, family = gaussian(), data = courtpenalties)
summary(model1)
```

We can extract information from the model with some generic helper functions: 

```{r}
# coefficients (including the intercept!)
coef(model1)

# coefficient confidence intervals
confint(model1)

# measures of model fit
BIC(model1) 

# residuals
head(residuals(model1))
```

# Tidy model output

Is there a tidier way? Yes! That's what `broom` is for.

Want tidy coefficients?

```{r}
tidy(model1)
```

Want a quick summary of model fit? 

```{r}
glance(model1)
```

Want fitted values and residuals?

```{r}
augment(model1)

augment(model1) %>%
  ggplot(aes(x = crt, y = .resid)) +
  geom_point() +
  labs(title = "Inspect residuals for heteroskedasticity")
```

# Coefficient plots

Coefficient plots are an alternative to looking at tables of coefficients.

```{r}
model1_df <- tidy(model1, conf.int = TRUE, conf.level = .95) 

model1_df %>%
  filter(term != "(Intercept)") %>%
  ggplot(aes(x = fct_rev(term), y = estimate)) + 
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip() + 
  labs(title = "Model 1: linear regression", 
       subtitle = "Estimated coefficient + 95% CI",
       x = NULL, 
       y = "coefficient")
```

These aren't standardized, so the scales are really different. 

Note: `ggstance::geom_pointrangeh()` also gives a horizontal version of `geom_pointrange()`. I felt like I was already using enough packages, though...

# Marginal effects plots

Instead of plotting coefficients, we can use those coefficients to get predictions on the y scale at different x values. In particular, we can plot specific *counterfactual* values of x. 

This is something we could do manually using the `predict()` function, but doing those calculations can be tricky with GLMs. (It's relatively easy with `lm()`.)

Instead, we'll use the `ggeffects` package. The `margins` package, which is based off of a Stata command, does something similar.

There are two ways of showing uncertainty in these plots:

- through the **confidence intervals** of the conditional means. This incorporates the uncertainty in the *systematic* part of the model---the estimates of your coefficients.
- through the **prediction intervals** of the data. This incorporates both the systematic uncertainty from the coefficients, and the uncertainty from the *stochastic* part of the model.

First, we'll pick one covariate, `crt`. By default, `ggpredict` holds the other continuous variables constant at their mean (and categorical variables at their reference values). 

```{r}
ggpredict(model1, terms = "crt")
```

```{r}
ggpredict(model1, terms = "crt", interval = "confidence", ci.level = .95) %>%
  plot() +
  labs(subtitle = "95% confidence interval")
```

```{r}
ggpredict(model1, terms = "crt", interval = "prediction", ci.lvl = .95) %>% 
  plot() +
  labs(subtitle = "95% prediction interval")
```

# Interaction terms

That's a simple association. Let's a second model, adding an interaction term:

```{r}
f2 <- update(f1, ~ . + crt*minority)
model2 <- glm(f2, family = gaussian(), data = courtpenalties)
tidy(model2)
```

Now the association between `crt` and `penalties` depends on the size of the non-white population. Let's pick 3 values of `minority` for our plot.

```{r}
model2_predictions <- ggpredict(model2, terms = c("crt", "minority [meansd]")) 

plot(model2_predictions) + theme(legend.position = "bottom")
```

Again, with prediction intervals instead: 

```{r}
ggpredict(model2, 
          terms = c("crt", "minority [meansd]"), 
          interval = "prediction") %>% 
  plot() + 
  labs(subtitle = "At mean non-white pop ± 1sd, 95% prediction interval") +
  theme(legend.position = "bottom")
```

Still want tables?

```{r message=FALSE, results='asis'}
library(texreg) # or stargazer, or pander ...
htmlreg(list(model1, model2), doctype = FALSE)
```

# Logistic regression

```{r}
grants <- read_csv("data/fedgrants&researchclimate.csv")

grants <- 
  grants %>%
  mutate(grant = ifelse(fedgrants > 0, 1, 0))

model3 <- glm(grant ~ hidegree + budget + climate, 
              family = binomial(link = "logit"),
              data = grants)

ggpredict(model3, terms = "budget [all]", interval = "confidence") %>% plot()
ggpredict(model3, terms = "budget [all]", interval = "prediction") %>% plot()
```

# Resources 

http://socviz.co/modeling.html

https://clanfear.github.io/CSSS508/Lectures/Week10/CSSS508_Week10_reproducibility_and_model_results.html

https://strengejacke.github.io/ggeffects/
